---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# conformalForecast <img src="man/figures/logo.svg" alt="logo" align="right" width="150" style="border: none; float: right;"/>

<!-- badges: start -->
[![R-CMD-check](https://github.com/xqnwang/conformalForecast/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/xqnwang/conformalForecast/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/conformalForecast)](https://CRAN.R-project.org/package=conformalForecast)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/grand-total/conformalForecast)](https://cran.r-project.org/package=conformalForecast)
[![License: GPL-3](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://github.com/xqnwang/conformalForecast/blob/main/LICENSE.md)
<!-- badges: end -->

The R package *conformalForecast* provides methods and tools for performing multistep-ahead time series forecasting using conformal prediction methods including classical conformal prediction, adaptive conformal prediction, conformal PID control, and autocorrelated multistep-ahead conformal prediction.

## Installation

You can install the development version of conformalForecast from [GitHub](https://github.com/xqnwang/conformalForecast) with:

``` r
# install.packages("remotes")
remotes::install_github("xqnwang/conformalForecast")
```

You can also get the official release version from [CRAN](https://CRAN.R-project.org/package=conformalForecast) with:

``` r
install.packages("conformalForecast")
```

## Example

This basic example shows how to perform classical conformal prediction and
update the results when new observations become available.

We simulate a series from an AR(2) model, holding back the last five
observations to stand in for data that arrives later:

```{r data, warning = FALSE}
library(conformalForecast)
set.seed(0)
series_all <- arima.sim(n = 1005, list(ar = c(0.8, -0.5)), sd = 1)
series <- head(series_all, 1000)
new_data <- as.numeric(tail(series_all, 5))
```

Conformal calibration needs a record of past forecast errors. We obtain one with
`cvforecast()`, which refits the model on a rolling forecast origin and stores
the out-of-sample errors at every horizon:

```{r cv}
far2 <- function(x, h, level) {
  Arima(x, order = c(2, 0, 0)) |>
    forecast(h = h, level = level)
}

fc <- cvforecast(series, forecastfun = far2, h = 3, level = c(80, 95),
                 forward = TRUE, initial = 1, window = 100)
```

Now calibrate intervals with split conformal prediction, on a rolling window of
100 calibration scores:

```{r scp}
scpfc <- conformal(fc, method = "scp", ncal = 100, rolling = TRUE)
```

When the five held-back observations arrive, `update()` extends the fit
incrementally. Intervals that were already issued do not change:

```{r update}
scpfc_updated <- update(scpfc, new_data = new_data, forecastfun = far2)
```

Finally, evaluate the intervals by forecast horizon:

```{r evaluate}
# Interval forecast accuracy
accuracy(scpfc_updated, byhorizon = TRUE)

# Mean coverage
coverage(scpfc_updated, level = 95)

# Mean and median interval width
width(scpfc_updated, level = 95, includemedian = TRUE)
```

## Learn more

- [Introduction to conformalForecast](https://xqnwang.github.io/conformalForecast/articles/conformalForecast.html): conformal methods, their formulas and tuning arguments, with a comparison of
  coverage and interval width.
- [Updating conformal forecasts with new data](https://xqnwang.github.io/conformalForecast/articles/update.html): how `update()` extends a fitted pipeline as observations arrive, and when it
  is equivalent to refitting.

## License
This package is free and open source software, licensed under GPL-3.
