---
title: "Introduction to conformalForecast"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to conformalForecast}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

The *conformalForecast* package provides some commonly used conformal prediction methods for time series forecasting.

```{r setup}
library(conformalForecast)
library(forecast)
library(ggplot2)
library(dplyr)
library(tibble)
library(tsibble)
```

## Data simulation

Suppose we are interested in forecasting a time series generated from an AR(2) model with $\phi_1 = 0.8$, $\phi_2=-0.5$, and $\sigma^2 = 1$.

```{r data, fig.width = 7, fig.height = 3}
set.seed(0)
series_all <- arima.sim(n = 1005, list(ar = c(0.8, -0.5)), sd = sqrt(1))
series <- head(series_all, 1000)
new_data <- as.numeric(tail(series_all, 5))
autoplot(series) +
  labs(
    title = "Time series generated from an AR(2) model"
  ) +
  theme_bw()
```

## Time series cross-validation

We first train a forecasting model AR(2) on a rolling forecast origin to generate forecasts and forecast errors on validation sets.

Setting `forward = TRUE` includes an $h$-step forecast from the final observation. The `window` argument controls the number of recent observations used at each forecast origin; setting it to `NULL` uses an expanding window.

```{r cv}
far2 <- function(x, h, level) {
  Arima(x, order = c(2, 0, 0)) |> forecast(h = h, level)
}
fc <- cvforecast(series, forecastfun = far2, h = 3, level = c(80, 95),
                 forward = TRUE, window = 100, initial = 1)
summary(fc)
```

```{r cvplot, fig.width = 7, fig.height = 3}
fc |>
  autoplot() +
  labs(
    title = "Forecasts produced using an AR(2) model"
  ) +
  theme_bw()
```

```{r cvinfo}
(fc_score <- accuracy(fc, byhorizon = TRUE))
(fc_cov <- coverage(fc, window = 100, level = 95))
(fc_wid <- width(fc, window = 100, level = 95, includemedian = TRUE))
```

## Conformal prediction

Based on the forecast errors on validation sets, we can apply various conformal prediction methods to construct prediction intervals.

Each method can be called directly or through `conformal()` by setting `method` to `"scp"`, `"acp"`, `"pid"`, or `"acmcp"`.

For SCP, ACP, and PID, `ncal` sets the calibration or burn-in length. If `rolling = TRUE`, only the most recent `ncal` errors are used where applicable; otherwise, the window expands over time. Setting `symmetric = TRUE` calibrates absolute errors, while `symmetric = FALSE` calibrates the lower and upper tails separately. AcMCP only supports asymmetric errors.

### Classical split conformal prediction (SCP)

Here, we perform a SCP method with equal weights in sample quantile estimation.

```{r scp}
scpfc <- conformal(fc, method = "scp", symmetric = FALSE, ncal = 100,
                   rolling = TRUE, weightfun = NULL, kess = FALSE,
                   quantiletype = 1)

(scpfc_score <- accuracy(scpfc, byhorizon = TRUE))
(scpfc_cov <- coverage(scpfc, window = 100, level = 95))
(scpfc_wid <- width(scpfc, window = 100, level = 95, includemedian = TRUE))
```

The `scp()` function allows us to include non-equal weights for sample quantile estimation by passing a weight calculation function to the `weightfun` argument.

```{r scp-exp}
expweight <- function(n) 0.99^{n+1-(1:n)}
scpfc_exp <- scp(fc, symmetric = FALSE, ncal = 100, rolling = TRUE,
                 weightfun = expweight, kess = FALSE, quantiletype = 1)

(scpfc_exp_score <- accuracy(scpfc_exp, byhorizon = TRUE))
(scpfc_exp_cov <- coverage(scpfc_exp, window = 100, level = 95))
(scpfc_exp_wid <- width(scpfc_exp, window = 100, level = 95, includemedian = TRUE))
```

### Adaptive conformal prediction (ACP)

The ACP method updates $\alpha$ online to target the desired long-run empirical coverage.

```{r acp}
acpfc <- acp(fc, symmetric = FALSE, gamma = 0.005, ncal = 100, rolling = TRUE)

(acpfc_score <- accuracy(acpfc, byhorizon = TRUE))
(acpfc_cov <- coverage(acpfc, window = 100, level = 95))
(acpfc_wid <- width(acpfc, window = 100, level = 95, includemedian = TRUE))
```

### Conformal PID control (PID)

The PID method combines three modules (quantile tracking, error integration, and scorecasting) to make an iteration to produce a sequence of quantile estimates used in the prediction sets.

```{r pid-setup}
# PID setup
Tg <- 1000; delta <- 0.01
Csat <- 2 / pi * (ceiling(log(Tg) * delta) - 1 / log(Tg))
KI <- 2
lr <- 0.1
```

```{r pid-nsf}
# PID without scorecaster
pidfc_nsf <- pid(fc, symmetric = FALSE, ncal = 100, rolling = TRUE,
                 integrate = TRUE, scorecast = FALSE,
                 lr = lr, Tg = Tg, KI = KI, Csat = Csat)

(pidfc_nsf_score <- accuracy(pidfc_nsf, byhorizon = TRUE))
(pidfc_nsf_cov <- coverage(pidfc_nsf, window = 100, level = 95))
(pidfc_nsf_wid <- width(pidfc_nsf, window = 100, level = 95, includemedian = TRUE))
```

```{r pid}
# PID with a Naive method as the scorecaster
naivefun <- function(x, h) {
  naive(x) |> forecast(h = h)
}
pidfc <- pid(fc, symmetric = FALSE, ncal = 100, rolling = TRUE,
             integrate = TRUE, scorecast = TRUE, scorecastfun = naivefun,
             lr = lr, Tg = Tg, KI = KI, Csat = Csat)

(pidfc_score <- accuracy(pidfc, byhorizon = TRUE))
(pidfc_cov <- coverage(pidfc, window = 100, level = 95))
(pidfc_wid <- width(pidfc, window = 100, level = 95, includemedian = TRUE))
```

### Autocorrelated multistep-ahead conformal prediction (AcMCP)

Similar to the PID method, the AcMCP method also integrates three modules (P, I, and D) to form the final iteration. However, instead of performing conformal prediction for each individual forecast horizon $h$ separately, AcMCP employs a combination of an MA($h-1$) model and a linear regression model of $e_{t+h|t}$ on $e_{t+h-1|t},\dots,e_{t+1|t}$ as the scorecaster. This allows the AcMCP method to capture the relationship between the $h$-step ahead forecast error and past errors.

The MA scorecaster uses CSS-ML estimation by default, matching the reference implementation. For longer forecast horizons, `ma_method = "CSS"` may be faster, although it can produce different estimates. Scorecasts are constructed recursively, so the first scorecast for horizon $h$ becomes available at cross-validation error index $n_{\mathrm{cal}} + h(h-1)/2$.

```{r acmcp}
acmcpfc <- acmcp(fc, ncal = 100, rolling = TRUE, integrate = TRUE, scorecast = TRUE,
             lr = lr, KI = KI, Csat = Csat)

(acmcpfc_score <- accuracy(acmcpfc, byhorizon = TRUE))
(acmcpfc_cov <- coverage(acmcpfc, window = 100, level = 95))
(acmcpfc_wid <- width(acmcpfc, window = 100, level = 95, includemedian = TRUE))
```

## Updating with new observations

When new observations become available, `update()` extends the cross-validation forecasts and conformal intervals without recomputing the existing results. It also reuses the conformal settings stored in the fitted object.

```{r update}
scpfc_updated <- update(scpfc, new_data = new_data, forecastfun = far2)
length(scpfc_updated$x)
scpfc_updated$mean
```

## Forecasting with external regressors

If the forecasting model uses external regressors, `forecastfun` should accept `xreg` for the training period and `newxreg` for the forecast period. The following smaller example uses the first 300 observations of the simulated series and reserves the next five for updating.

```{r xreg}
far2_xreg <- function(x, h, level, xreg, newxreg) {
  Arima(x, order = c(2, 0, 0), xreg = xreg) |>
    forecast(h = h, level = level, xreg = newxreg)
}

n <- 300
h <- 3
n_update <- 5
series_xreg <- head(series_all, n)
new_data_xreg <- as.numeric(series_all[n + seq_len(n_update)])
xreg_all <- cbind(
  trend = seq_len(n + h + n_update),
  cycle = sin(2 * pi * seq_len(n + h + n_update) / 12)
)

fc_xreg <- cvforecast(
  series_xreg,
  forecastfun = far2_xreg,
  h = h,
  level = 95,
  forward = TRUE,
  window = 100,
  xreg = xreg_all[seq_len(n + h), ]
)
scpfc_xreg <- conformal(
  fc_xreg,
  method = "scp",
  symmetric = FALSE,
  ncal = 100,
  rolling = TRUE
)
```

With `forward = TRUE`, the stored `xreg` already contains the original $h$ future rows. Therefore, `new_xreg` supplies the rows immediately after the stored predictor matrix, keeping the next forecast horizon available after the update. It must have one row per new observation and the same columns as the stored `xreg`.

```{r xreg-update}
scpfc_xreg_updated <- update(
  scpfc_xreg,
  new_data = new_data_xreg,
  forecastfun = far2_xreg,
  new_xreg = xreg_all[n + h + seq_len(n_update), ]
)
scpfc_xreg_updated$mean
```

## Coverage and width of prediction intervals

Taking the AcMCP result as an example, we now show the average coverage on validation sets.

```{r covplot, fig.width = 7, fig.height = 5}
acmcpfc_cov$rollmean |>
  as_tsibble() |>
  mutate(horizon = key, coverage = value) |>
  update_tsibble(key = horizon) |>
  select(-c(key, value)) |>
  ggplot(aes(x = index, y = coverage, group = horizon)) +
  geom_line() +
  geom_hline(yintercept = 0.95, linetype = "dashed", color = "blue") +
  facet_grid(horizon~., scales = "free_y") +
  xlab("Time") +
  ylab("Rolling mean coverage for AcMCP") +
  theme_bw()
```

We can also show the rolling average interval width on validation sets.

```{r widplot, fig.width = 7, fig.height = 5}
acmcpfc_wid$rollmean |>
  as_tsibble() |>
  mutate(horizon = key, width = value) |>
  update_tsibble(key = horizon) |>
  select(-c(key, value)) |>
  ggplot(aes(x = index, y = width, group = horizon)) +
  geom_line() +
  facet_grid(horizon~., scales = "free_y") +
  xlab("Time") +
  ylab("Rolling mean width for AcMCP") +
  theme_bw()
```

We can also combine all the results and show them in one single plot. Here, WCP denotes weighted SCP with exponential weights, and PI denotes conformal PID without the scorecasting component. The comparison is based on a single simulated series and is intended only to illustrate the output; it should not be used to rank the methods.

```{r bind-cov, fig.width = 7, fig.height = 5}
candidates <- c("fc", "scpfc", "scpfc_exp", "acpfc", "pidfc_nsf", "pidfc", "acmcpfc")
methods <- c("AR", "SCP", "WCP", "ACP", "PI", "PID", "AcMCP")
for (i in 1:length(candidates)) {
  out <- get(paste0(candidates[i], "_cov"))
  out_pivot <- out$rollmean |>
    as_tsibble() |>
    mutate(horizon = key, coverage = value) |>
    update_tsibble(key = horizon) |>
    select(-c(key, value)) |>
    mutate(method = methods[i]) |>
    as_tibble()
  assign(paste0(methods[i], "_cov"), out_pivot)
}
cov <- bind_rows(mget(paste0(methods, "_cov")))

cols <- c(
  "AR" = "black",
  "SCP" = "yellow",
  "WCP" = "#fa9200",
  "ACP" = "green",
  "PI" = "blue",
  "PID" = "purple",
  "AcMCP" = "red"
)
cov |>
  as_tsibble(index = index, key = c(horizon, method)) |>
  mutate(method = factor(method, levels = methods)) |>
  ggplot(aes(x = index, y = coverage, group = method, colour = method)) +
  geom_line(linewidth = 0.8, alpha = 0.8) +
  scale_colour_manual(values = cols) +
  geom_hline(yintercept = 0.95, linetype = "dashed", colour = "gray") +
  facet_grid(horizon~.) +
  xlab("Time") +
  ylab("Rolling mean coverage") +
  theme_bw()
```

```{r bind-covmean}
cov_mean <- lapply(1:length(candidates), function(i) {
  out_cov <- get(paste0(candidates[i], "_cov"))
  out_score <- get(paste0(candidates[i], "_score"))
  out_mean <- data.frame(
      method = methods[i],
      covmean = as.vector(out_cov$mean),
      winkler = as.vector(out_score[, "Winkler_95"]),
      msis = as.vector(out_score[,"MSIS_95"])
    ) |>
    as_tibble() |>
    rownames_to_column("horizon") |>
    mutate(horizon = paste0("h=", horizon))
  out_mean
})
cov_mean <- do.call(bind_rows, cov_mean) |>
  mutate(method = factor(method, levels = methods)) |>
  mutate(covdiff = covmean - 0.95) |>
  arrange(horizon, method)
print(cov_mean, n = nrow(cov_mean))
```

```{r bind-wid, fig.width = 7, fig.height = 5}
for (i in 1:length(candidates)) {
  out <- get(paste0(candidates[i], "_wid"))
  out_pivot <- out$rollmean |>
    as_tsibble() |>
    mutate(horizon = key, width = value) |>
    update_tsibble(key = horizon) |>
    select(-c(key, value)) |>
    mutate(method = methods[i]) |>
    as_tibble()
  assign(paste0(methods[i], "_wid"), out_pivot)
}
wid <- bind_rows(mget(paste0(methods, "_wid")))

wid |>
  as_tsibble(index = index, key = c(horizon, method)) |>
  mutate(method = factor(method, levels = methods)) |>
  ggplot(aes(x = index, y = width, group = method, colour = method)) +
  geom_line(linewidth = 0.8, alpha = 0.8) +
  scale_colour_manual(values = cols) +
  facet_grid(horizon~.) +
  xlab("Time") +
  ylab("Rolling mean width") +
  theme_bw()
```
